---
title: "One Way ANOVA"
date: "`03-14-2023`"
output:
  html_document:
    theme: default
    highlight: textmate
    toc: yes
    toc_float: yes
    fig_width: 9.67
    fig_height: 6
    fig_caption: yes
  pdf_document:
    toc: yes
---

# Two-way ANOVA
ANOVA test is used to compare multiple samples and derive inferences about their respective populations

# Assumptions
- Population variances are equal
- Populations are normal

# Important Note
To clear the workspace, use the command below:
```{r}
rm(list = ls())
```

# Summary Statistics

```{r}
I <- 4
J <- 3
K <- 2

data <- c(
            # b_1         #b_2          #b_3
            876, 913,     1413, 1290,   1291, 1412, #a_1
            1156, 1219,   1876, 1710,   2115, 1963, #a_2
            1234, 1181,   1591, 1649,   1650, 1712, #a_3
            825, 797,     1083, 1161,   1148, 1262  #a_4
            )

factor_a <- factor(
        c(
            1, 1, 1, 1, 1, 1,
            2, 2, 2, 2, 2, 2,
            3, 3, 3, 3, 3, 3,
            4, 4, 4, 4, 4, 4
            ))
factor_b <- factor(
        c(
            1, 1, 2, 2, 3, 3,
            1, 1, 2, 2, 3, 3,
            1, 1, 2, 2, 3, 3,
            1, 1, 2, 2, 3, 3
            ))
```



# Built-in ANOVA in R

## With Interaction

```{r}
# run 2-Way ANOVA with interaction
anova_with_interaction <- aov(data ~ factor_a * factor_b)
summary(anova_with_interaction)
```
## Without Interaction

```{r}
# run 2-Way ANOVA without interaction
anova_without_interaction <- aov(data ~ factor_a + factor_b)
summary(anova_without_interaction)
```

# Long Hand ANOVA Calculations

```{r}
# Sums of Squares

```

# Hypothesis Testing
```{r}

# Null hypothesis      : mu_1 = mu_2 = ... = mu_n
# Alternate hypothesis : at least one of group means are different

# Test Statistic

# look up P-value (verify with table)

```

# Data Scatter plot
```{r}

```