---
title: "Probability"
date: "`01-16-2023`"
output:
  html_document:
    theme: default
    highlight: textmate
    toc: yes
    toc_float: yes
    fig_width: 9.67
    fig_height: 6
    fig_caption: yes
  pdf_document:
    toc: yes
---

# Important Note
To clear the workspace, use the command below:
```{r}
rm(list = ls())
```


# Statistic Summary
```{r}
mu <- 3 # set the mean
sigma <- 2 # set the standard deviation
```


# Normal Distribution

## Area (pnorm)
Calculate the area under the normal distribution graph which is equivalent to finding the p-value
```{r}
# calculating areas (i.e. looking up p-values)
# for the N(mu,sigma^2), i.e. normal dist with mean=mu and std dev=sigma, look up area to the right of X=z (upper tail)
mu <- 0
sigma <- 1
z <- 1.36108931
isLowerTail <- FALSE
pnorm(z, mean = mu, sd = sigma, lower.tail = isLowerTail) * 2
```
## Critical Values (qnorm)
If area is known and we need to find the critical value for which the area to the left or right of it is equal to given area
```{r}
# finding critical values 
# for the N(mu,sigma^2), i.e. normal dist with mean=mu and std dev=sigma
# look up the value of X where the area to the right (the upper tail area) is "area"
mu <- 0
sigma <- 1
area <- 2
isLowerTail <- TRUE

qnorm(area, mean = mu, sd = sigma, lower.tail = isLowerTail)
```



# Student-t Distribution

## Area (pt)
Calculate the area under the student-t distribution
```{r}
# calculating areas (i.e. looking up p-values)
# for the t(dof) distribution, look up area to the left of X=t (lower tail)

dof <- 5
t <- 1
isLowerTail <- TRUE

pt(t, df = dof, lower.tail = isLowerTail)
```

## Critical value (qt)
```{r}
# finding critical values
# for the t(dof) distribution, look up the value of X where the area to the right (upper tail area) is "area"

dof <- 5
area <- 0.1
isLowerTail <- FALSE

qt(area, df = dof, lower.tail = isLowerTail)
```









